Fail Faster

### 2025-03-22 paper/paper.tex
work on intro

We evaluate the optimizations with a series of case studies in each
generator library, and show that the \name{} generators run faster than their unstaged equivalents.
Further, we show generators that run faster also find bugs faster
% ability \jwc{The paper is called ``fail faster'' --- we shouldn't say ``bug finding ability'', we should say something like ``how fast you can squash bugs''. This is too passive }
by running these case studies in the Etna
platform~\cite{etna}, which simulates real-world PBT usage \iflater\hg{Don't love this phrase, I'll
think more about how we should talk about Etna}\fi by injecting bugs into a system and
measuring how quickly various generators detect them.

In summary, we show that PBT generator DSLs incur significant performance costs across languages, and present a general technique for improving their efficiency while preserving their idiomatic abstractions. \jwc{needs a bit about RNG --- BCP thoughts?}
We make the following contributions:
\begin{enumerate}
    \item We identify two key sources of inefficiency in PBT generator libraries, which can significantly impact performance:
    abstraction overhead and \rand choice.
    \item We present \name{}, a technique for eliminates the abstraction overhead of generators. We implement \name{} in both OCaml and Scala 3, showcasing its cross-language applicability.
    \item We demonstrate that programs written using \name{} yield performance improvements to generator running time. Finally, we show
    that these performance improvements extend to significantly improved bug-finding speed. \jwc{This should be sell-ier: 2x speedups, etc. ``bug-finding speed'' isn't a race against time.}
    \jwc{We precisely quantify the performance effects of staging and faster rng across a range of testing scenarios --- i.e. a contrib is the fact that we can exactly pin down when one helps versus the other. The ``logically smarter'' your generator is, the more you get out of staging, the ``dumber'', the more you get out of rng.}
    \bcp{Agree with this. We need to beef up our claims here about what we learned about PBT-specifics.}
    \jwc{And this needs to be flowed through the rest of the paper.}
\end{enumerate}

>>>>

We evaluate the optimizations with a series of case studies in each
generator library, and show that the \name{} generators run faster than their unstaged equivalents.
Further, we show generators that run faster also find bugs faster
by running these case studies in the Etna
platform~\cite{etna}, which simulates real-world PBT usage \iflater\hg{Don't love this phrase, I'll
think more about how we should talk about Etna}\fi by injecting bugs into a system and
measuring how quickly various generators detect them.

In summary, we show that PBT generator DSLs incur significant performance costs across languages, and present a general technique for improving their efficiency while preserving their idiomatic abstractions. \jwc{needs a bit about RNG --- BCP thoughts?}
We make the following contributions:
\begin{enumerate}
    \item We identify two key sources of inefficiency in PBT generator libraries, which can significantly impact performance:
    abstraction overhead and \rand choice.
    \item We present \name{}, a technique for eliminates the abstraction overhead of generators. We implement \name{} in both OCaml and Scala 3, showcasing its cross-language applicability.
    \item We demonstrate that writing generators using \name{} and fast \rands{} yields performance improvements.
    We precisely quantify the impact of each optimization separately, and explain how they play distinct yet complementary roles in improving generator performance.
    Finally, we show that these performance improvements extend to significantly improved bug-finding speed, finding some bugs as much as \todo{\%\%} faster.
\end{enumerate}
